Raft Distributed Consensus Algorithm over WebRTC (Web Real-Time Communication)

### 2015-09-16 research_class/paper.tex
Plural -> singular for subject/verb felicity.
Directionalize all opening double-quotes.

In order to ensure that each state machine on every node executes
exactly the same commands in exactly the same order, the Raft system
provides safety guarantees. In particular, there are some restrictions
on whether a Raft node is actually eligible to become a leader, based on
the state of its transaction log compared to other node transaction
logs. A Raft node will only vote for a candidate if the candidate has
a log that is more up-to-date than the voter. A log entry with a later
term is always more up-to-date. If the log entries have the same term,
then the entry with a higher index is more up-to-date. In addition,
Raft leaders may not consider entries from a previous term to be
committed until it has committed at least one entry from its own
term
\cite[Section~3.6]{raft_thesis:ongaro14}.

@@

The modular design of the Raft.js implementation allows it to be
easily used in different contexts. For example, the \texttt{RaftServerLocal}
class (\texttt{local.js}) implements RPC calls using plain JavaScript function
calls between different instances of the class (nodes) in the same
execution context. For example, the online Raft visualization at
\href{http://kanaka.github.io/raft.js/}{http://kanaka.github.io/raft.js/}
\cite{raft.js:visualization} uses the \texttt{RaftServerLocal} class
and instantiates it using a scheduling function that executes one
scheduled event each time the user clicks on the "Step" button rather
than due to the passage of time.

@@

The Raft over WebRTC testing application is designed as a single web
page (\texttt{rtc.html} or \texttt{chat.html}). The application has
different starting behavior depending on the value of the URL query
string (the URL component following the "?" and up to the "\#") and
the fragment identifier (the URL component after the "\#").

@@

If the URL fragment identifier is set to "firstServer", then this Raft
node is initialized as the leader of a single node cluster and all
normal timers are started. The fragment identifier is then unset. The
query parameter contains the ID of this cluster (the PeerJS channel).

Subsequent nodes are added to the same Raft cluster by starting a new
browser context (separate browser, browser window, or browser tab) and
loading the same URL without the "firstServer" fragment identifier
(Figure~\ref{fig:raft_rtc_sequence}~[9]). There is a convenience link
provided in the application page that opens a new browser window with
a new cluster node (see \ref{fig:browser_1}).

@@

When an instance of the web application starts that is not
a "firstServer", the normal Raft timers are not enabled until the
server receives its first RPC. The prevents the server from
trying to becoming a leader before it is incorporated into the cluster
by an appendEntries RPC from the true cluster leader.

@@

\begin{enumerate}
\item the \texttt{clientRequestResponse} is registered with the local
    Raft.js implementation as the callback function anytime
    a \texttt{clientRequestResponse} RPC is received.
\item the \texttt{clientRequest} function keeps track of the current
    leader ID.  If the current leader ID is not set or this node is
    the leader then the client request RPC is sent directly to the
    local Raft node bypassing the network transport layer.  If the
    current leader ID is set then the request is sent to that node as
    a normal RPC request over the WebRTC channel.
\item when the \texttt{clientRequestResponse} function is called and
    the status is "NOT\_LEADER" then the current leader ID being
    tracked by the \texttt{clientRequest} function is set to the
    leader ID specified in the response. Then the original request is
    re-issued over the normal WebRTC channel to the leader.
\item when the \texttt{clientRequestResponse} function is called with
    a successful status, the original callback argument of the
    \texttt{clientRequest} function is called with the results from
    the RPC response.
\end{enumerate}

@@

Then four more Raft nodes were then added to the cluster by clicking
on the embedded "create more nodes" link in the application.
Figure~\ref{fig:browser_5} shows a screenshot of the browser once four
new nodes are added to the cluster (5 total):

@@

Then four more Raft nodes were then added to the cluster by clicking
on the embedded "create more clients" link in the application.
Finally, on a non-leader node more text lines were entered and sent by
clicking the Send button. Figure~\ref{fig:browser_chat_5} shows
a screenshot of the browser once four new nodes/tabs are added to the
cluster (5 total) and another text line was sent from the fifth node:

@@

To simplify the execution of the tests and to make the testing
infrastructure more portable, a docker container environment was used
to encapsulate and execute the SlimerJS environment. The container is
built using \emph{test/Dockerfile} which is in turn derived from
the "fentas/slimerjs" docker image \cite{fentas:slimerjs}.
This custom Dockerfile was created because the upstream version of the
Dockerfile used an older version of XULRunner that did not support
more than 7 WebRTC clients communicating together simultaneously. This
issue was discovered during testing.

@@

Figure 7 is a scatterplot with each test run as a blue "x" and a black
line representing the average at each cluster size.

@@

\begin{enumerate}
    \item \emph{Figure 8 (a)}: The results show that once the cluster
        size reaches 17, the cluster sometimes does not activate
        before the test times out (the red "x" marks). Once the
        cluster reaches 21 nodes, all 7 test runs time out. The reason
        for this is because at 17 nodes, the leader is not always able
        to send out appendEntries RPCs to all the other nodes before
        their election timeout fires. More analysis is needed, but one
        possible explanation is that the single SlimerJS XULRunner
        process that is handling all the cluster nodes is not able to
        give enough CPU time to the leader node to reliable send
        appendEntries RPCs to all the other nodes within the election
        timeout value. Increasing the election timeout to 500ms or
        greater eliminates the failures without significantly changing
        the average cluster activation time curve.
    \item \emph{Figure 9 (a)}: The results show that when the cluster
        size reaches 15, then average amount of time for cluster
        recovery (after the leader is killed) increases dramatically
        (not all data points are shown on the plot for size 15). At
        cluster size 17 the cluster is no longer able to recover
        within the test timeout period. Increasing the election
        timeout to 500ms or greater eliminates the failures at the
        cost of slightly increased recovery time for smaller cluster
        sizes.
    \item \emph{Figure 10 (a)}: During this test cluster was unable to
        activate at cluster size 21 when the election timeout value is
        set to 100ms (see Figure 8 (a)). This means there was no
        opportunity for cluster recovery at cluster size 21 so there
        is no data for that column.

@@

Each figure is a scatterplot with each test run as a blue "x" and
a black line representing the average at each cluster size.

>>>>

In order to ensure that each state machine on every node executes
exactly the same commands in exactly the same order, the Raft system
provides safety guarantees. In particular, there are some restrictions
on whether a Raft node is actually eligible to become a leader, based on
the state of its transaction log compared to other node transaction
logs. A Raft node will only vote for a candidate if the candidate has
a log that is more up-to-date than the voter. A log entry with a later
term is always more up-to-date. If the log entries have the same term,
then the entry with a higher index is more up-to-date. In addition,
a Raft leader may not consider entries from a previous term to be
committed until it has committed at least one entry from its own
term
\cite[Section~3.6]{raft_thesis:ongaro14}.

@@

The modular design of the Raft.js implementation allows it to be
easily used in different contexts. For example, the \texttt{RaftServerLocal}
class (\texttt{local.js}) implements RPC calls using plain JavaScript function
calls between different instances of the class (nodes) in the same
execution context. For example, the online Raft visualization at
\href{http://kanaka.github.io/raft.js/}{http://kanaka.github.io/raft.js/}
\cite{raft.js:visualization} uses the \texttt{RaftServerLocal} class
and instantiates it using a scheduling function that executes one
scheduled event each time the user clicks on the ``Step" button rather
than due to the passage of time.

@@

The Raft over WebRTC testing application is designed as a single web
page (\texttt{rtc.html} or \texttt{chat.html}). The application has
different starting behavior depending on the value of the URL query
string (the URL component following the ``?" and up to the ``\#") and
the fragment identifier (the URL component after the ``\#").

@@

If the URL fragment identifier is set to ``firstServer", then this Raft
node is initialized as the leader of a single node cluster and all
normal timers are started. The fragment identifier is then unset. The
query parameter contains the ID of this cluster (the PeerJS channel).

Subsequent nodes are added to the same Raft cluster by starting a new
browser context (separate browser, browser window, or browser tab) and
loading the same URL without the ``firstServer" fragment identifier
(Figure~\ref{fig:raft_rtc_sequence}~[9]). There is a convenience link
provided in the application page that opens a new browser window with
a new cluster node (see \ref{fig:browser_1}).

@@

When an instance of the web application starts that is not
a ``firstServer", the normal Raft timers are not enabled until the
server receives its first RPC. The prevents the server from
trying to becoming a leader before it is incorporated into the cluster
by an appendEntries RPC from the true cluster leader.

@@

\begin{enumerate}
\item the \texttt{clientRequestResponse} is registered with the local
    Raft.js implementation as the callback function anytime
    a \texttt{clientRequestResponse} RPC is received.
\item the \texttt{clientRequest} function keeps track of the current
    leader ID.  If the current leader ID is not set or this node is
    the leader then the client request RPC is sent directly to the
    local Raft node bypassing the network transport layer.  If the
    current leader ID is set then the request is sent to that node as
    a normal RPC request over the WebRTC channel.
\item when the \texttt{clientRequestResponse} function is called and
    the status is ``NOT\_LEADER" then the current leader ID being
    tracked by the \texttt{clientRequest} function is set to the
    leader ID specified in the response. Then the original request is
    re-issued over the normal WebRTC channel to the leader.
\item when the \texttt{clientRequestResponse} function is called with
    a successful status, the original callback argument of the
    \texttt{clientRequest} function is called with the results from
    the RPC response.
\end{enumerate}

@@

Then four more Raft nodes were then added to the cluster by clicking
on the embedded ``create more nodes" link in the application.
Figure~\ref{fig:browser_5} shows a screenshot of the browser once four
new nodes are added to the cluster (5 total):

@@

Then four more Raft nodes were then added to the cluster by clicking
on the embedded ``create more clients" link in the application.
Finally, on a non-leader node more text lines were entered and sent by
clicking the Send button. Figure~\ref{fig:browser_chat_5} shows
a screenshot of the browser once four new nodes/tabs are added to the
cluster (5 total) and another text line was sent from the fifth node:

@@

To simplify the execution of the tests and to make the testing
infrastructure more portable, a docker container environment was used
to encapsulate and execute the SlimerJS environment. The container is
built using \emph{test/Dockerfile} which is in turn derived from
the ``fentas/slimerjs" docker image \cite{fentas:slimerjs}.
This custom Dockerfile was created because the upstream version of the
Dockerfile used an older version of XULRunner that did not support
more than 7 WebRTC clients communicating together simultaneously. This
issue was discovered during testing.

@@

Figure 7 is a scatterplot with each test run as a blue ``x" and a black
line representing the average at each cluster size.

@@

\begin{enumerate}
    \item \emph{Figure 8 (a)}: The results show that once the cluster
        size reaches 17, the cluster sometimes does not activate
        before the test times out (the red ``x" marks). Once the
        cluster reaches 21 nodes, all 7 test runs time out. The reason
        for this is because at 17 nodes, the leader is not always able
        to send out appendEntries RPCs to all the other nodes before
        their election timeout fires. More analysis is needed, but one
        possible explanation is that the single SlimerJS XULRunner
        process that is handling all the cluster nodes is not able to
        give enough CPU time to the leader node to reliable send
        appendEntries RPCs to all the other nodes within the election
        timeout value. Increasing the election timeout to 500ms or
        greater eliminates the failures without significantly changing
        the average cluster activation time curve.
    \item \emph{Figure 9 (a)}: The results show that when the cluster
        size reaches 15, then average amount of time for cluster
        recovery (after the leader is killed) increases dramatically
        (not all data points are shown on the plot for size 15). At
        cluster size 17 the cluster is no longer able to recover
        within the test timeout period. Increasing the election
        timeout to 500ms or greater eliminates the failures at the
        cost of slightly increased recovery time for smaller cluster
        sizes.
    \item \emph{Figure 10 (a)}: During this test cluster was unable to
        activate at cluster size 21 when the election timeout value is
        set to 100ms (see Figure 8 (a)). This means there was no
        opportunity for cluster recovery at cluster size 21 so there
        is no data for that column.

@@

Each figure is a scatterplot with each test run as a blue ``x" and
a black line representing the average at each cluster size.
